\textbf{Proposal for bachelor thesis}\\ Exploration of the influence of graph parameters on the generalization error of Graph Neural Networks

### 2024-09-02 paper/main.tex
final proposal version v1

\documentclass{article}
\usepackage[margin=1.5in]{geometry}

@@

\begin{abstract}
The goal of the bachelor thesis is to understand empirically how graph parameters/characteristics influence the generalization error of Graph Neural Networks (GNNs).
\end{abstract}

\section{Setup}
In following, I will descripe training framework, datasets and models for the setup of the experiment.

@@

\subsection{Models}
I have used GCN layer~\cite{kipf2017semisupervised} and GAT layer~\cite{velickovic2020pointer} in the experiment. In the futher experiment, I will consider to use other GNN layers. In the Appendix, I will provide the code of the example models.

@@

\begin{figure}[h]
    \includegraphics[width=0.5\textwidth]{experiment_procedure.pdf}
    \caption{A illutration of experiment Procedure}
    \label{fig:experiment}
\end{figure}

@@

@@

\begin{table}[]
    \centering
    \begin{tabular}{@{}lll@{}}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Dataset}} & \multicolumn{2}{c}{Parameter}                                            \\ \cmidrule(l){2-3} 
    \multicolumn{1}{c}{}                         & \multicolumn{1}{c}{Ave. degree} & \multicolumn{1}{c}{Ave. shortest path} \\ \midrule
    FRANKENSTEIN                                 & 2.0627         & 3.6521                     \\
    NCI1                                         & 2.1550         & 5.4697                     \\
    COIL-RAG                                     & 1.8277         & 1.0606                     \\
    Letter-high                                  & 1.8896         & 1.5173                     \\
    DD                                           & 4.9790         & 7.9700                     \\
    PTOTEIN\_full                                & 3.7346         & 4.7123                     \\
    COLORS-3                                     & 2.9288         & 2.0945                     \\ \bottomrule
    \end{tabular}
    \caption{The calculated parameters from a set of datasets}
    \label{tab:parameter}
\end{table}

@@

class GCN(torch.nn.Module):
    def __init__(self, in_channels, hidden_channels, out_channels):
        super().__init__()
        self.conv1 = GCNConv(in_channels, hidden_channels)
        self.conv2 = GCNConv(hidden_channels, hidden_channels)
        self.conv3 = GCNConv(hidden_channels, hidden_channels)
        self.conv4 = GCNConv(hidden_channels, hidden_channels)
        self.conv5 = GCNConv(hidden_channels, hidden_channels)
        self.fc1 = nn.Linear(hidden_channels, hidden_channels)
        self.fc2 = nn.Linear(hidden_channels, out_channels)

    def forward(self, x, edge_index, batch, edge_weight=None):
        x = self.conv1(x, edge_index, edge_weight).relu()
        x = self.conv2(x, edge_index, edge_weight).relu()
        x = self.conv3(x, edge_index, edge_weight).relu()
        x = self.conv4(x, edge_index, edge_weight).relu()
        x = self.conv5(x, edge_index, edge_weight).relu()
        x = global_mean_pool(x, batch)
        x = self.fc1(x).relu()
        x = F.dropout(x, p=0.5, training=self.training)
        x = self.fc2(x)
        return x

>>>>

\documentclass{article}
\usepackage{geometry}
\geometry{
    a4paper,
    left=30mm,
    right=30mm,
    top=20mm,
    bottom=20mm
}

@@

\begin{abstract}
The goal of this bachelor thesis is to understand empirically how graph parameters/characteristics influence the generalization error of Graph Neural Networks (GNNs).
\end{abstract}

\section{Introduction}

What is graph parameters? Graph parameters are the properties of the graph, which can be calculated from the graph. For example, the average degree of the graph, the average shortest path of the graph, the average number of coloring in the 1-WL algorithm, etc. In the early experiment, the graph parameters are calculated as the average degree and the average shortest path of the graphs.
\\
In the another hand, the generalization error is a measure of how well the model generalizes to unseen data. In the experiment, the generalization error is calculated as the difference between the training accuracy and the test accuracy.
\\
Why is it important to understand the influence of graph parameters on the generalization error of GNNs? The understanding of the influence of graph parameters on the generalization error of GNNs can help us to understand the behavior of GNNs. For example, if we know that the average degree of the graph has a significant influence on the generalization error of GNNs, we can design GNNs that are more robust to the average degree of the graph.

\section{Setup in the early experiment}
In following, I will descripe training framework, datasets and models for the setup of the experiment.

@@

\subsection{Models}
In the early experiment, I have used GCN layer~\cite{kipf2017semisupervised} and GAT layer~\cite{velickovic2020pointer}. To prevent overfitting, dropout is used in the model. Relu is used as activation function. In the Appendix, I have provided the code of example models.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{experiment_procedure.pdf}
    \caption{A illutration of experiment Procedure}
    \label{fig:experiment}
\end{figure}

@@

\subsection{Experimental details in the early stage}
I use Adam optimizer with learning rate 0.01. The batch size is set to 64. The hidden dimension of the GNN layer is set to 64. The number of epochs is set to 200. To prevent overfitting and to save time, the early stopping is used with patience 20. The loss function is set to CrossEntropyLoss.

\section{Results}

@@

\begin{table}[h!]
    \centering
    \begin{tabular}{@{}lll@{}}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Dataset}} & \multicolumn{2}{c}{Parameter}                                            \\ \cmidrule(l){2-3} 
    \multicolumn{1}{c}{}                         & \multicolumn{1}{c}{Ave. degree} & \multicolumn{1}{c}{Ave. shortest path} \\ \midrule
    Mutagenicity                                 & 2.0379         & 4.4825                     \\
    NCI1                                         & 2.1550         & 5.4697                     \\
    COIL-RAG                                     & 1.8277         & 1.0606                     \\
    Letter-high                                  & 1.8896         & 1.5173                     \\
    DD                                           & 4.9790         & 7.9700                     \\
    PTOTEIN\_full                                & 3.7346         & 4.7123                     \\
    COLORS-3                                     & 2.9288         & 2.0945                     \\ \bottomrule
    \end{tabular}
    \caption{The calculated parameters from a set of datasets}
    \label{tab:parameter}
\end{table}

\begin{table}[h!]
    \centering
    \begin{tabular}{@{}lll@{}}
    \toprule
    Dataset        & Ave. generalization error & Standard Deviation   \\ \midrule
    Mutagenicity   & 0.0268      & 0.02335 \\
    NCI1           & 0.0138      & 0.02649 \\
    COIL-RAG       & 0.0450      & 0.01378 \\
    Letter-high    & 0.0403      & 0.04342 \\
    DD             & 0.0524      & 0.05849 \\
    PROTEINS\_full & 0.0189      & 0.05032 \\
    COLORS-3       & 0.0076      & 0.01428 \\ \bottomrule
    \end{tabular}
    \caption{The calculated generalization error from a set of datasets}
    \label{tab:error}
\end{table}

Then I plot the generalization error against the graph parameters. The result is shown in Figure~\ref{fig:scatter}.

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{final_results.png}
    \caption{The generalization error against the graph parameters}
    \label{fig:scatter}
\end{figure}

\section{In the further work}
In the future, I plan to explore different GNN layers and datasets to evaluate the consistency of the results. Additionally, I will consider experimenting with various graph parameters. In \cite{gera2018annotated}, the authors have proposed a set of graph parameters that can be used to characterize the graph. I will use these parameters in the future work.

@@

class GCN(torch.nn.Module):
    def __init__(self, in_channels, hidden_channels, out_channels):
        super().__init__()
        self.conv1 = GCNConv(in_channels, hidden_channels)
        self.conv2 = GCNConv(hidden_channels, hidden_channels)
        self.conv3 = GCNConv(hidden_channels, hidden_channels)
        self.conv4 = GCNConv(hidden_channels, hidden_channels)
        self.fc1 = nn.Linear(hidden_channels, hidden_channels)
        self.fc2 = nn.Linear(hidden_channels, out_channels)

    def forward(self, x, edge_index, batch, edge_weight=None):
        x = self.conv1(x, edge_index, edge_weight).relu()
        x = self.conv2(x, edge_index, edge_weight).relu()
        x = self.conv3(x, edge_index, edge_weight).relu()
        x = self.conv4(x, edge_index, edge_weight).relu()
        x = global_mean_pool(x, batch)
        x = self.fc1(x).relu()
        x = F.dropout(x, p=0.5, training=self.training)
        x = self.fc2(x)
        return x
